---
author: "Thomas Wutzler"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    keep_md: true
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Processing several measurement cycles of different setup}
  %\usepackage[UTF-8]{inputenc}
---

```{r setup, include = FALSE}
# rmarkdown::render("vignettes/switchingChambers.Rmd")
library(knitr)
opts_chunk$set(
    #out.extra = 'style = "display:block; margin: auto"'
    #, fig.align = "center"
    fig.width = 4.3, fig.height = 3.2, dev.args = list(pointsize = 10)
    , message = FALSE
    , results = 'hold'
    )
knit_hooks$set(spar = function(before, options, envir) {
    if (before) {
        par( las = 1 )                   #also y axis labels horizontal
        par(mar = c(2.0,3.3,0,0) + 0.3 )  #margins
        par(tck = 0.02 )                          #axe-tick length inside plots             
        par(mgp = c(1.1,0.2,0) )  #positioning of axis title, axis labels, axis
     }
})
```


Processing several measurement cycles of different setup
========================================================

```{r eval=FALSE, echo=FALSE, results='hide'}
isDevelopMode <- TRUE
library(twDev)
setwd('..');loadPkg()
```
```{r results = 'hide', warning = FALSE}
if (!exists("isDevelopMode")) library(RespChamberProc)
set.seed(0815)      # for reproducible results
```

Determine subsets of single measurement cycles
----------------------------------------------

First, the data is loaded. Here, directly from zipped logger-output.
```{r concAllPlot, spar = TRUE, fig.width = 12}
library(dplyr)
# fit chambers in parallel inside calcClosedChamberFluxForChunkSpecs
library(furrr)
plan(multisession, workers = 4) 

fName <- system.file(
  "genData/SMANIE_Chamber1_26032015.zip", package = "RespChamberProc")
if (nzchar(fName) ) { 
  ds <- readDat(
    unz(fName, filename = unzip(fName, list = TRUE)[1,"Name"] ),tz = "UTC") }
head(ds)
plot( CO2_LI840 ~ TIMESTAMP, ds, ylab = "CO2 (ppm)", xlab = "Time")
```

The dataset contains several measurement cycles of light and dark chambers 
with increasing or decreasing concentrations respectively.

First, we correct the pressure to standard units and correct the CO2
concentrations for water vapour.
```{r}
ds$Pa <- ds$AirPres * 100  # convert hPa to Pa
ds$CO2_dry <- corrConcDilution(ds, colConc = "CO2_LI840", colVapour = "H2O_LI840")
ds$H2O_dry <- corrConcDilution(ds, colConc = "H2O_LI840", colVapour = "H2O_LI840")
ds$VPD <- calcVPD( ds$SurTemp, ds$Pa, ds$H2O_LI840)
```

In order to process each measurement cycle independently, we first determine 
parts of the time series that are contiguous, i.e. without gaps and without 
change of an index variable, here variable `collar`.

```{r}
dsChunk <- subsetContiguous(ds, colTime = "TIMESTAMP", colIndex = "Collar") 
head(dsChunk)
```

The modified dataset contains a new variable, `iChunk`, which reports a factor that
changes with different measurement cycles.
This factor can be used to select subset of single measurement cycles. 
```{r concSinglePlot, spar = TRUE}
dss <- subset(dsChunk, iChunk == 15)
plot( CO2_dry ~ TIMESTAMP, dss, ylab = "CO2 (ppm)", xlab = "time (Minute:Second)")
```

Associating Chamber information to chunks
-----------------------------------------
Different collars may have different depth leading to different volume, or are
connected by tubing of different length to the sensor.

The user needs to provide 
- a mapping of iChunk to collar in column `collar`
- dimensional information about the collar in a data.frame

Function `subsetContiguous` creates a default for the collar mapping
by assigning the values of the index column.

```{r}
dsChunk %>% group_by(iChunk) %>% summarise(collar = first(collar)) %>%  head()
```

DataFrame `collar_spec` then needs to specify for each collar id in column `collar`, 
the colums `area` (m2) and `volume` (m3),
as well a `tlag` (s), the lag time between start of the cycle
, i.e. the start of the chunk (usually chamber closing time), 
and the time when the gas reaches the sensor.

In this example, we specify the same surface area and the same tlag 
for each collar but simulate 
removing the litter and then measuring the depth of each collar 
to recompute the volume. The depth, here, are random numbers around 3 cm.
If the lagtime is set to missing (NA) then it is estimated 
in each chunk by a breakpoint detection.


```{r}
chamberVol = 0.6*0.6*0.6		# chamber was a cube of 0.6m length
surfaceArea = 0.6*0.6
collar_spec <- tibble(
  collar = unique(dsChunk$collar), 
  depth = pmax(0,rnorm(length(collar), mean = 0.03, sd = 0.015)),
  area = surfaceArea,
  volume = chamberVol + surfaceArea * depth,
  tlag = NA)
head(collar_spec)
```

Problems with association setups to the data can be checked by function
`checkCollarSpec`, which returns FALSE and attribute `msg` in case of problems.
```{r results='hide'}
checkCollarSpec(dsChunk, collar_spec)
```


Computing the flux
------------------

Function `calcClosedChamberFluxForChunkSpecs` applies 
function `calcClosedChamberFlux` to each subset.

```{r}
# for demonstration use only the first 20 cycles
dsChunk20 <- subset(dsChunk, as.integer(iChunk) <= 20) 
resChunks1 <- calcClosedChamberFluxForChunkSpecs(
  dsChunk20, collar_spec
  , colTemp = "AirTemp"
  # linear and saturating shape
  , fRegress = c(lin = regressFluxLinear, tanh = regressFluxTanh)	
  , debugInfo = list(omitEstimateLeverage = TRUE)	# faster
)
head(resChunks1)
```

The results are similar as for `calcClosedChamberFlux`, unless there are 
several rows identified by additional key columns `iChunk` and `chamber.`

## Plotting faceted data and fits
We recommend to plot the results together with the concentration data
to dectect problems. 
```{r fittedPlots, spar = TRUE, fig.width = 11, fig.height = 6, warning = FALSE, message = FALSE}
library(ggplot2)
plots <- plotCampaignConcSeries( dsChunk20, resChunks1, plotsPerPage = 64L)  
print(plots$plot[[1]]) # print the first page
```

If argument `fileName` is provided to `plotCampaignConcSeries`. All plots are
written to a pdf. If there are more cycles, i.e. plots, than argument 
`plotsPerPage` (default 64) there will be several pages in the pdf.

## Inspecting lag-times

Lag times between closing the chamber and the start of the concentration
increase, i.e. when the gas arrives at the sensor, is by default estimated by a
breakpoint detection method. This method is not robust to fluctuations, early
saturation, or other possible peculiarities of the concentration time series. In
other to detect those subsets, where lag-time detection has failed, on can
inspect the inferred lag-times for outliers.

For a campaign where all the measurement cycles were performed with similar
conditions, the lag-time should not differ much.

```{r inspectLagTimes}
table(resChunks1$tLag)
```

The plots do not indicate problems, and the longest estimated lagtime
varies between 0 and 16 seconds. 
Since slightly overestimating the lagtime does not change the flux but
only might slightly increase the uncertainty, 
we infer that for this campaign a lag-time of about 16 seconds is appropriate.

One can save processing time and avoid failures in the non-robust
breakpoint-detection by specifying
a fixed lag-time (may differ across collars) with the collar specification.
```{r}
collar_spec2 <- mutate(collar_spec, tlag = 16)

resChunks2 <- calcClosedChamberFluxForChunkSpecs(
  dsChunk20, collar_spec2
  , colTemp = "AirTemp"
  # linear and saturating shape
  , fRegress = c(lin = regressFluxLinear, tanh = regressFluxTanh)	
  , debugInfo = list(omitEstimateLeverage = TRUE)	# faster
)
head(resChunks2)
```
```{r eval=FALSE, include=FALSE}
ds2 <- resChunks2 %>% 
  select(iChunk, flux) %>% 
  rename( fluxFixed = flux )
ds <- left_join(resChunks1, ds2, by = "iChunk")
#ggplot( ds, aes(flux, fluxFixed)) + geom_point()
ggplot( ds, aes(flux, I((fluxFixed - flux)/flux)*100) ) + geom_point()
```



